---
title: "Project 1_Alexa"
author: "Alexa Owen"
date: "2023-06-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
Data <- read.csv("diamonds4.csv")
library (tidyverse)
library(cowplot)
```

```{r Clarity vs. Cut}
ggplot((Data) , aes(x=clarity, fill=cut))+
  geom_bar(position = "fill")+
  scale_x_discrete(limits=c("SI2", "SI1", "VS2", "VS1", "VVS2", "VVS1", "IF", "FL")) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray"))  +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  labs(x="Diamond Clarity", y="Proportion", 
       title="Proportion of Diamond Cuts by Diamond Clarity")

clarity_cut <- table (Data$clarity, Data$cut)
clarity_cut

prop_clarity_cut <- round(prop.table(clarity_cut, 1) * 100, 2)
prop_clarity_cut
```
The website claims that many diamond characteristics can be hidden by cut. To test this claim, diamond cut was analyzed against diamond clarity. From Blue Nile's claim, the quality of the diamond cut should be higher in diamonds with lower clarity. According to the website, the clarity ratings from lowest to highest in the data set are SI2, SI1, VVS2, VVS1, VS2, VS1, IF, FL. Furthermore, the ratings can be grouped together as SI, VS, VVS, IF and FL. Additionally, the cut quality from low to high is Good, Very Good, Ideal, Astor Ideal. Based on these rankings and the claim, we should expect to see lower clarity ratings (SI and VS) associated with higher cut quality. From the bar chart we can see that the FL clarity rating only contained the diamond cuts Very Good and Ideal. The second highest clarity rating contained Good, Very Good, and Ideal diamond cuts. The clarity ratings SI2 to VS1 contained diamonds with all four cuts.

From the bar chart this claim appears to be supported by the data. For the purpose of looking at this data the Astor Ideal cut is going to be ignored in the discussion as there were not many diamond purchases with this cut giving very little data. For the SI2 clarity, 52.12% of the cuts were Ideal which is the second highest cut ranking overall. Except for the Astor Ideal cut, the Good cut had the lowest percentage at 8.48%. The SI1 clarity showed similar results with, 61.73% of the cuts being Idea and the cut of Good had the lowest percentage at 7.82%. For VS2, the Ideal cut was highest at 61.21% and Good was lowest at 6.07. Finally, the VS1 cut had the highest percentage of Ideal cuts at 58.37% and the Good cut was the lowest at 5.58%. From these proportions, the claim that better diamond cuts hide imperfections in the diamond appears to be true. 


```{r Carat vs. Price}

carat_.5 <- Data %>% 
  filter(between(carat,.4,.6))

carat_1 <- Data %>% 
  filter(between(carat,.9,1.1))

carat_1.5 <- Data %>% 
  filter(between(carat,1.4,1.6))

carat_2 <- Data %>% 
  filter(between(carat,1.9,2.1))

graph_.5<- ggplot((carat_.5), aes(x=carat, y=price))+
  geom_point(alpha = .5, color = "blue")+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray")) +
  labs(x="Carat (0.40-0.60)", y="Price", 
       title="Diamond Carat Against Price")

graph_1<- ggplot((carat_1), aes(x=carat, y=price))+
  geom_point(alpha = .5, color = "blue")+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray")) +
  labs(x="Carat (0.90-1.10)", y="Price", 
       title="Diamond Carat Against Price")

graph_1.5<- ggplot((carat_1.5), aes(x=carat, y=price))+
  geom_point(alpha = .5, color = "blue")+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray")) +
  labs(x="Carat (1.40-1.60)", y="Price", 
       title="Diamond Carat Against Price")

graph_2<- ggplot((carat_2), aes(x=carat, y=price))+
  geom_point(alpha = .5, color = "blue")+
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray"))  +
  labs(x="Carat (1.90-2.10)", y="Price", 
       title="Diamond Carat Against Price")

plot_grid(graph_.5, graph_1, graph_1.5, graph_2, labes = "AUTO", nrow= 2, ncol = 2)
```

The claim Blue Nile makes is that diamond carat weights that are slightly less (0.4, 0.9, 1.4, 1.9) than typical carat values (0.5, 1.0, 1.5, or 2.0) cost less than getting a normal carat weight. From the data this claim is best supported by lower carat weights such as a half of a carat and a carat. The top two graphs show that a diamond with a carat weight of 0.4 does appear to have a lower cost than diamonds with a carat weight of 0.5. This is further displayed by the second graph which shows a similar trend. The carat weight of 0.9 appears to have a lower price than a carat weight of 1.0. As the carat value continues to increase it becomes more difficult to make assumptions from the graph because there are fewer data points. Therefore for diamonds with a carat value greater than 1.0, it is difficult to see if the data actually supports Blue Nile's claim. As a result, we can say that we can assume their claim appears to be supported for carat values of 1.0 or less, but is inconclusive at higher carat weights. 

```{r Cut drives price}
ggplot((Data) , aes(x=cut, y=price, color = cut))+
  geom_boxplot ()+
  coord_cartesian(ylim=c(0, 20000)) +
  scale_x_discrete(limits=c("Good", "Very Good", "Ideal", "Astor Ideal")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray"))  +
  labs(x="Diamond Cut", y="Price", 
       title="Diamond Cut Against Diamond Price")

ggplot((Data) , aes(x=clarity, y=price, color = clarity))+
  geom_boxplot ()+
  coord_cartesian(ylim=c(0, 30000)) +
  scale_x_discrete(limits=c("SI2", "SI1", "VVS2", "VVS1", "VS2", "VS1", "IF", "FL")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray"))  +
  labs(x="Diamond Clarity", y="Price", 
       title="Diamond Clarity Against Diamond Price")

ggplot((Data) , aes(x=color, y=price, color = color))+
  geom_boxplot ()+
  coord_cartesian(ylim=c(0, 20000)) +
  scale_x_discrete(limits=c("J", "I", "H", "G", "F", "E", "D")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray"))  +
  labs(x="Diamond Color", y="Price", 
       title="Diamond Color Against Diamond Price")

ggplot(Data, aes(x = xstar, y = ystar)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  geom_point(color = "blue4", alpha = 0.7, size = 2)+
  geom_smooth(method=lm,color="darkorange",alpha=0.5,se = FALSE)+
  labs(title = "Price Distribution by Carat",
       x = "Carat",
       y = "Price") +
  scale_y_continuous(labels = scales::comma, breaks = seq(0, max(Data$price), by = 50000)) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray")) 

```

```{r Median by Cut}
aggregate(Data$price,        
          list(Data$cut),
          median)

aggregate(Data$price,        
          list(Data$clarity),
          median)

aggregate(Data$price,        
          list(Data$color),
          median)

cor (Data$price, Data$carat)
```

The main claim made by Blue Nile is that diamond cut drives the price of the diamond. If this claim is true, we should expect to see an increase in median prices as the cut quality increases and the other variables of carat, color, and clarity should not show as much difference as quality or weight increases. From the box plot that compares diamond cut and price, the claim does not appear to be true. As stated, if cut drove price, then a diamond cut of Good should have the lowest median price while better diamond cuts should have higher median prices. In reality, what is shown is that as cut quality increases, except for Astor Ideal, the median diamond price decreases. It should be noted that Astor Ideal is Blue Nile's exclusive cut which could impact the price. For a diamond cut of Good, the median diamond cost was 1,903 dollars, for Very Good it was 1,744.5 dollars, and for Ideal it was 1354.0 dollars, and for Astor Ideal it was 2854.0 dollars. In an attempt to see what characteristics drive the cost, box plots were also created for diamond clarity and color against price and a scatter plot was created for carat and price. The box plot for clarity shows that except for Flawless (FL) diamonds, the median price of the diamonds is relatively similar from low clarity to high clarity. The box plot for diamond color showed a similar result with most medians being close to 1,500 dollars. The scatter plot for carat against price showed a positive r value of 0.82. This suggests that as carat weight increases, price does as well. From these four graphs combined, it appears that we can debunk Blue Nile's claim that cut drives price. Instead, it appears that the diamond's carat weight is what drive the cost of the diamond. 
