---
title: "Project 1"
author: "Fadumo Hussein, Karolina Straznikiewicz, Alexa Owen, and Kaleigh O’Hara"
date: "2023-06-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dbplyr)
library(ggplot2)
library(tidyverse)
Data = read.csv(" ")
```

Data 

This dataset contains information about one thousand diamonds available for sale on Blue Nile (http://www.bluenile.com), a renowned online diamond retailer. The data set encompasses various attributes of the diamonds, including carat weight, clarity, color, cut, and price. The data was collected from Blue Nile's website, capturing a diverse range of diamonds offered by the company.

Blue Nile is a prominent online jewelry retailer specializing in diamonds and engagement rings. Established in 1999, it has become a trusted destination for customers seeking high-quality diamonds at competitive prices. Blue Nile offers a wide selection of certified diamonds sourced from reputable suppliers around the world. 

Please note that while this data set primarily originates from Blue Nile, additional internet resources were utilized to complete it.

Variables 

1. Carat 

Carat weight refers to the measurement of a diamond's mass, with one carat equivalent to 200 milligrams or 0.2 grams, and is often used as an indicator of a diamond's size or perceived value.

- numeric (in grams)

2. Clarity 

Clarity of a diamond refers to the presence or absence of internal and external flaws. It is evaluated on a scale of various levels of clarity grades, impacting the diamond's overall visual appearance and value. Diamonds surface flaws are called blemishes, while internal defects are known as inclusions.

- categorical 

Levels: 
- "FL"  are Flawless (FL) Diamonds - no internal or external characteristics - very rare
- "IF" are Internally Flawless (IF) Diamonds - have no inclusions within the stone
- "VVS1" and "VVS2" are Very, Very Slightly Included (VVS) Diamonds 
- "VS1" and "VS2" are  Very Slightly Included (VS) Diamonds
- "SI1" and "SI2" are Slightly Included (SI) Diamonds  

3. Color 

Color of a diamond refers to the presence or absence of any noticeable tint or hue within the stone. Diamond color is graded on a scale from D (colorless) to Z (light yellow or brown), with higher grades indicating a more colorless and desirable diamond.

- categorical 

Levels: "D" - Absolutely Colorless, "E" - Colorless, "F" - Colorless, "G" - Near Colorless,
"H" - Near Colorless (faint yellow hue), "I" - Near Colorless (slight yellow), 
"J" - Near Colorless (light yellow tone), "K" - Faint

4. Cut 

Cut of a diamond refers to the quality of how the diamond was shaped. A well-cut diamond reflects light very well, resulting in enhanced sparkle and a higher value, while a poorly cut diamond may appear dull and less captivating.

- categorical

Levels:
- "Astor Ideal" - Cut from the finest raw material, crafted to gather and reflect the most light possible.

- "Ideal" - 3% of diamond cut quality. It reflects most of the light that enters the diamond.

- "Very Good" - 15% of diamond cut quality. It reflects nearly as much light as the ideal cut.

- "Good" - 25% of diamond cut quality. It reflects most of the light that enters, but not as much as a Very Good cut grade.

5. Price 

The price of diamond. 

- numeric (in dollars)


```{r}
Data$clarity = as.factor(Data$clarity)
Data$color = as.factor(Data$color)
Data$cut = as.factor(Data$cut)

Data$clarity<-factor(Data$clarity, 
                     levels=c("SI1","SI2","VS1","VS2","VVS1","VVS2", "IF","FL"))

#levels(Data$clarity)

Data$cut<-factor(Data$cut, 
                     levels=c("Good", "Very Good", "Ideal","Astor Ideal"))

#levels(Data$cut)

Data$color<-factor(Data$color, 
                     levels=c("K","J","I","H","G","F","E","D"))

#levels(Data$color)
```

Graphs: 

Clarity - Price 

```{r}
ggplot(Data, aes(x = price, y = clarity)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "none") +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.5, fill = "blue4", alpha = 0.7) +
  labs(title = "Price by Clarity",
       x = "Price",
       y = "Clarity") +
  scale_x_continuous(labels = scales::comma) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray"))

```


```{r}
library(ggplot2)
library(ggthemes)

ggplot(Data, aes(x = clarity, y = price)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  geom_point(color = "blue4", alpha = 0.7, size = 2)+
  labs(title = "Price Distribution by Clarity",
       x = "Clarity",
       y = "Price") +
  scale_y_continuous(labels = scales::comma, breaks = seq(0, max(Data$price), by = 50000)) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray")) 

```


```{r}
ggplot(Data, aes(x = clarity, y = price)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12, face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  #geom_boxplot(fill = "blue4", alpha = 0.7, width = 0.5) +
  geom_violin(fill = "blue4", alpha = 0.7, width = 0.5) +
  labs(title = "Price Distribution by Clarity",
       x = "Clarity",
       y = "Price") +
  scale_y_continuous(labels = scales::comma) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray"))

```


```{r}
avg_price_by_clarity <- Data %>%
  group_by(clarity) %>%
  summarise(avg_price = mean(price))

ggplot(avg_price_by_clarity, aes(x = clarity, y = avg_price)) +
  geom_bar(stat = "identity", fill = "blue4", alpha = 0.7) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "none") +
  labs(title = "Average Price by Clarity",
       x = "Clarity",
       y = "Average Price") +
  scale_y_continuous(labels = scales::comma) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray"))

```


Price vs Carat

```{r}
ggplot(Data, aes(x = carat, y = price)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  geom_point(color = "blue4", alpha = 0.7, size = 2)+
  geom_smooth(method=lm,color="darkorange",alpha=0.5,se = FALSE)+
  labs(title = "Price Distribution by Carat",
       x = "Carat",
       y = "Price") +
  scale_y_continuous(labels = scales::comma, breaks = seq(0, max(Data$price), by = 50000)) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray")) 
```
Price vs Cut

```{r}
ggplot(Data, aes(x = cut, y = price)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  geom_point(color = "blue4", alpha = 0.7, size = 2)+
  labs(title = "Price Distribution by Cut",
       x = "Cut",
       y = "Price") +
  scale_y_continuous(labels = scales::comma, breaks = seq(0, max(Data$price), by = 50000)) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray")) 
```

```{r}

avg_price_by_clarity2 <- Data %>%
  group_by(cut) %>%
  summarise(avg_price = mean(price))

ggplot(avg_price_by_clarity2, aes(x = cut, y = avg_price)) +
  geom_bar(stat = "identity", fill = "blue4", alpha = 0.7) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "none") +
  labs(title = "Average Price by Cut",
       x = "Cut",
       y = "Average Price") +
  scale_y_continuous(labels = scales::comma) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray"))

```

Price vs Color

```{r}
ggplot(Data, aes(x = color, y = price)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12,face = "bold"),
        legend.title = element_blank(),
        legend.position = "none") +
  geom_point(color = "blue4", alpha = 0.7, size = 2)+
  labs(title = "Price Distribution by Color",
       x = "Color",
       y = "Price") +
  scale_y_continuous(labels = scales::comma, breaks = seq(0, max(Data$price), by = 50000)) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray")) 
```

```{r}
avg_price_by_clarity3 <- Data %>%
  group_by(color) %>%
  summarise(avg_price = mean(price))

ggplot(avg_price_by_clarity3, aes(x = color, y = avg_price)) +
  geom_bar(stat = "identity", fill = "blue4", alpha = 0.7) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "none") +
  labs(title = "Average Price by Color",
       x = "Color",
       y = "Average Price") +
  scale_y_continuous(labels = scales::comma) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray"))
```

```{r}
ggplot(Data, aes(x = price, y = color)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        axis.text = element_text(size = 10, face = "bold"),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "none") +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.5, fill = "blue4", alpha = 0.7) +
  labs(title = "Price by Color",
       x = "Price",
       y = "Color") +
  scale_x_continuous(labels = scales::comma) +
  theme(plot.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "lightgray"))

```
Carat vs. Clarity

```{r}
Data <- read.csv("diamonds4.csv")
summary(Data$carat)
Data <- Data %>%
  mutate(carat.cat = cut(carat, breaks =c(-Inf, .4, .52, 1, Inf), labels = c("1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile")))
clarity_levels  = factor(Data$clarity, levels=c("SI1", "SI2", "VS1", "VS2", "VVS1", "VVS2", "IF", "FL"))
Data <- data.frame(Data, clarity_levels)

ggplot(subset(data, !is.na(carat.cat)), aes(x=clarity_levels, fill=carat.cat))+
  geom_bar(position = "fill")+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x = "Clarity", y = "Proportion of Diamonds", title = "Diamond Clarity vs. Diamond Carat")

ggplot(subset(data, !is.na(carat.cat)), aes(x=carat.cat, fill=clarity_levels))+
  geom_bar(position = "fill")+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x = "Clarity", y = "Proportion of Diamonds", title = "Diamond Clarity vs. Diamond Carat")
```
The first plot shows that of diamonds with the highest clarity, Flawless (FL), over 50% were in the highest carat quartile as well. This supports the webstie's claim that as you go up in carat weight, other characteristics become more visible. However, in all other categories, there is no distinct difference in the distribution of carat and level of clarity. This leads to the hypothesis that for a diamond's carat to have an effect on its clarity, the diamond needs to be over 1 carat. However, since this dataset is of diamonds that exist on the market, perhaps diamonds of an extremely high carat also have extremely high clarity because the clientele for a high carat is also the clientele for a high clarity.

Cut vs. Color
```{r}

Data$color<-factor(Data$color, 
                     levels=c( "D",   "E",   "F",   "G",   "H",  "I", "J"))

Data$cut<-factor(Data$cut, 
                     levels=c("Good", "Very Good", "Ideal","Astor Ideal"))

Data %>%
   ggplot(aes(x = cut, fill = color))+ 
   geom_bar(position = "fill") +
    theme(axis.text.x = element_text(angle = 0))+
    labs(x="Type of cut", y="Percentage of Color in Each Cut", 
         title="Distribution of color in each cut")

```

Color vs. Clarity
```{r}
Data$color<-factor(Data$color, 
                     levels=c( "D",   "E",   "F",   "G",   "H",  "I", "J"))


Data$clarity<-factor(Data$clarity, 
                     levels=c("SI1","SI2","VS1","VS2","VVS1","VVS2", "IF","FL"))
Data %>%
   ggplot(aes(x = color, fill = clarity))+ 
   geom_bar(position = "fill") +
    theme(axis.text.x = element_text(angle = 0))+
    labs(x="Color", y="Percentage of Clarity in Each Color",
         title="Distribution of Clarity for each Color")

```
